Magnet analysis

### 2023-06-07 paper/manuscript.Rmd
clean up analysis files

# Introduction
- why do we care about developmental trajectory? ref to stat learning paper
- variation

@@

- [@raviv2018developmental]
- [@astor2022gaze]
- [@colombo2001development]
- [@scaife1975capacity]
- [@itakura1998use]

@@

- [@silverstein2021infants] for vertical plane
- [@zhang2019role]
- [@frischen2007gaze]
- [@lee1998children]
- [@coelho2006searching]

@@

# Lifespan 
- development & individual differences in gaze understanding
- verweis methods paper reliable differences kinder & adults. 
- kontinuerliche, systematische variation, wodrin? => model

@@

```{r fig1, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab1)"}
knitr::include_graphics("../figures/tango_procedure.png")
```

@@

# Computational cognitive model
- model auch für warum ältere Erwachsene schlecht? anders schlecht als kinder?
- entwicklungsperspektive basierend auf model
- dann prozess-ebene => magnet

@@

@@

The authors made the following contributions:
### TODO

>>>>

# Introduction

-   why do we care about developmental trajectory? ref to stat learning paper
-   variation

@@

-   [@raviv2018developmental]

-   [@astor2022gaze]

-   [@colombo2001development]

-   [@scaife1975capacity]

-   [@itakura1998use]

@@

-   [@silverstein2021infants] for vertical plane

-   [@zhang2019role]

-   [@frischen2007gaze]

-   [@lee1998children]

-   [@coelho2006searching]

@@

# Lifespan

-   development & individual differences in gaze understanding
-   verweis methods paper reliable differences kinder & adults.
-   kontinuerliche, systematische variation, wodrin? =\> model

<https://osf.io/6yjz3>

@@

```{r fig1, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab1)"}
knitr::include_graphics("../figures/magnet_procedure.png")
```

@@

# Computational cognitive model

-   model auch für warum ältere Erwachsene schlecht? anders schlecht als kinder?
-   entwicklungsperspektive basierend auf model
-   dann prozess-ebene =\> magnet

@@

Task design, data collection, and sample sizes were pre-registered: <https://osf.io/xsqkt>.

The study design and procedure obtained ethical clearance by the MPG Ethics commission Munich, Germany, falling under a packaged ethics application (Appl. No. 2021_45), and was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology.
The research adheres to the legal requirements of psychological research with children in Germany.
Data were collected between February and March 2023.

-   gut, dass unser maß ind diff messen kann, das nutzen wir jetzt

-   zsmhang mit vector following, aber nicht redundant

-   perspective-taking auch, aber auch nicht relevant, nicht theory of mind score

-   diskutieren dass die maße auch nict ind diff maße sind

-   model vergleich: welche komponente stecken in gafo drin?,
    welchen beitrag liefern die um score zu erklären?
    welche wichtiger, tom oder nur perspective taking

-   magnet versuch das vector following zu isolieren.
    ist da noch was anderes außerhalb vector-following?

-   aggregierte messungen immer besser, nur indikatoren für konstrukt

@@

The authors made the following contributions: \### TODO

### 2023-06-07 stats/magnet.Rmd
clean up analysis files

---
title: "Stats"
author: "Julia Prein"
date: "03/21/2023"
output: pdf_document
---

@@

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(corrr)
library(psych)
library(apaTables)

@@

```{r load_data}
magnet <- readRDS("../data/magnet-data.rds")
```

```{r scatterplots}
magnet %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 5, label.y = 6, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")

magnet %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 3.5, label.y = 3.25, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  scale_y_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")
```

```{r corrplot}
magnet %>% 
  select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% 
  rename("Tango mean" = tango_mean, "Magnet mean" = magnet_mean, "ToM aggregate score" = tom_aggregate, "Perspective-taking score" = perspective_aggregate) %>% 
  cor() %>% 
  corrplot(diag = FALSE, type = 'upper', method = 'ellipse', order = 'FPC', col = COL2('BrBG'), tl.col = 'black', cl.ratio = 0.2, tl.srt = 30, addCoef.col ='black', number.cex = 0.8)
```

```{r cor_matrix}
tango_magnet <- magnet %>% select(tango_mean, magnet_mean) %>% corr.test(method = "pearson")
tango_tom <- magnet %>% select(tango_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")
magnet_tom <- magnet %>% select(magnet_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")

@@

# apa table function only supports pearson corr so far
magnet %>% select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% apa.cor.table(filename = "APA_Correlation_Table.doc", table.number=1)
```

```{r plot_cor_ci}
cor_matrix %>% 
  filter(Correlation != "ToM x Perspective-taking") %>% 
  mutate(Correlation = fct_relevel(Correlation, "Magnet x Perspective-taking", "Magnet x ToM", "Tango x Perspective-taking", "Tango x ToM", "Tango x Magnet")) %>%
  ggplot(., aes(x=Correlation, y=r, ymin=lower, ymax=upper)) +
  geom_pointrange(color = "#005555", size = 0.8, linewidth = 1) +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("") +
  coord_flip() +
  theme(axis.title.x = element_text(face="italic"))

ggsave("../figures/magnet_cor_ci.png", width = 5, height = 3)
```
```

>>>>

---
title: "Magnet analysis"
author: "Julia Prein"
date: "06/07/2023"
output: pdf_document
---

@@

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(psych) # corr.test
library(brms) # modeling
library(tidybayes) # modeling
library(ggpubr) # stat_cor 
library(corrplot) # corr plot
library(apaTables) # as name says

@@

```{r load_data}
magnet_data <- readRDS("../data/magnet-data.rds")
```

```{r sample}
magnet_sample <- magnet_data %>% 
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = sum(female, na.rm = T),
  )

magnet_sample

write.csv(magnet_sample, "../saves/magnet_sample.csv", quote = F, row.names = F)
saveRDS(magnet_sample, "../saves/magnet_sample.rds")
```

```{r scatterplots}
# doesnt make a difference whether we exclude outlier / "imprecise" kids
magnet_data %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 5, label.y = 6, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")

magnet_data %>% 
  ggplot(., aes(x = magnet_mean, y = tango_mean)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_smooth(method = "lm", size = 1, alpha = 0.25, col = "darkgrey") +
  geom_point(color = "#005555") +
  stat_cor(method = "pearson", label.x = 3.5, label.y = 3.25, aes(label = paste(..r.label..)), size = 5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  scale_y_continuous(limits = c(0, 3.75), breaks = seq(0, 4, 0.5)) +
  labs(x = "Magnet mean", y = "Tango mean")
```

```{r corrplot}
magnet_data %>% 
  select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% 
  rename("Tango mean" = tango_mean, "Magnet mean" = magnet_mean, "ToM score" = tom_aggregate, "Perspective-taking score" = perspective_aggregate) %>% 
  cor() %>% 
  corrplot(diag = FALSE, type = 'lower', method = 'color', order = 'FPC', col = COL2('BrBG'), tl.col = 'black', cl.ratio = 0.2, tl.srt = 0, addCoef.col ='black', number.cex = 0.8)
```

```{r cor_matrix}
tango_magnet <- magnet_data %>% select(tango_mean, magnet_mean) %>% corr.test(method = "pearson")
tango_tom <- magnet_data %>% select(tango_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")
magnet_tom <- magnet_data %>% select(magnet_mean, tom_aggregate, perspective_aggregate) %>% corr.test(method = "spearman")

@@

# apa table function only supports pearson corr so far
magnet_data %>% select(tango_mean, magnet_mean, tom_aggregate, perspective_aggregate) %>% apa.cor.table(filename = "../saves/magnet_corrtable.doc", table.number=1)
```

```{r plot_cor_ci}
cor_ci <- cor_matrix %>% 
  filter(Correlation != "ToM x Perspective-taking") %>% 
  mutate(Correlation = fct_relevel(Correlation, "Magnet x Perspective-taking", "Magnet x ToM", "Tango x Perspective-taking", "Tango x ToM", "Tango x Magnet")) %>%
  ggplot(., aes(x=Correlation, y=r, ymin=lower, ymax=upper)) +
  geom_pointrange(color = "#005555", size = 0.8, linewidth = 1) +
  scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("") +
  coord_flip() +
  theme(axis.title.x = element_text(face="italic"))

cor_ci

saveRDS(cor_ci, file = "../saves/magnet_cor_ci.rds")
# ggsave("../figures/magnet_cor_ci.png", width = 5, height = 3)
```

```{r check_tango_distribution}
ggplot(magnet_data, aes(x = tango_mean)) + 
  geom_density() + 
  stat_function(fun = dnorm, 
                args = with(magnet_data, c(mean = mean(tango_mean), sd = sd(tango_mean))), 
                color = "red", linetype = "dotted") +

  stat_function(fun = dlnorm, 
                args = with(magnet_data, c(meanlog = mean(log(tango_mean)), sdlog = sd(log(tango_mean)))),
                color = "blue", linetype = "dotted")
```

```{r prepare_modeldata}
mdata <- magnet_data %>% 
  mutate(
    magnet_scaled = scale(magnet_mean), 
    tom_scaled = scale(tom_aggregate), 
    perspective_scaled = scale(perspective_aggregate)
  ) %>% 
  select(subj_id, tango_mean, age_centered, magnet_scaled, tom_scaled, perspective_scaled)
```

```{r models, message=FALSE, warning=FALSE}
mage <- brm(tango_mean ~ age_centered,
                    data = mdata,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4) %>%
  add_criterion(c("loo","waic"))

mmagnet <- update(mage, formula. = ~ . + magnet_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mtom <- update(mage, formula. = ~ . + tom_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mperspective <- update(mage, formula. = ~ . + perspective_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mmagnet_tom <- update(mage, formula. = ~ . + magnet_scaled + tom_scaled, newdata = mdata) %>% 
  add_criterion(c("loo","waic"))

mmagnet_perspective <- update(mage, formula. = ~ . + magnet_scaled + perspective_scaled, newdata = mdata) %>%
  add_criterion(c("loo","waic"))
```

```{r modelcomparison}
mcomparison <- loo_compare(mage, mmagnet, mtom, mperspective, mmagnet_tom, mmagnet_perspective, criterion = "waic") %>% as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mage, mmagnet, mtom, mperspective, mmagnet_tom, mmagnet_perspective, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mcomparison

model_results <- mcomparison %>%
  select(model, waic, se_waic, modelweight, elpd_diff, se_diff) %>%
  rename(
    Model = model, 
    WAIC = waic, 
    SE_WAIC = se_waic,
    Weight = modelweight, 
    ELPD_DIFF = elpd_diff, 
    SE_ELPD_DIFF = se_diff,
  ) %>% 
  mutate(
    Model = recode(Model, 
                   "mage" = "Null model with Age (scaled)", 
                   "mmagnet" = "Magnet mean (scaled)", 
                   "mtom" = "ToM aggregate (scaled)", 
                   "mperspective" = "Perspective-taking aggregate (scaled)", 
                   "mmagnet_tom" = "Magnet mean (scaled) + ToM aggregate (scaled)", 
                   "mmagnet_perspective" = "Magnet mean (scaled) + Perspective-taking aggregate (scaled)", 
                   )
  )

write.csv(model_results, "../saves/magnet_modelresults.csv", quote = F, row.names = F)
saveRDS(model_results, "../saves/magnet_modelresults.rds")
```

```{r magnet_draws}
magnet_draws <-mmagnet_perspective %>% 
  gather_draws(b_age_centered, b_magnet_scaled, b_perspective_scaled) %>% 
  mutate(
    .variable = factor(.variable, 
                       levels = c("b_age_centered", "b_perspective_scaled", "b_magnet_scaled"), 
                       labels = c("Age (scaled)", "Perspective-taking (scaled)", "Magnet mean (scaled)")), 
  ) %>% 
  rename(estimate = .value, predictor = .variable)

write.csv(magnet_draws, "../saves/magnet_draws.csv", quote = F, row.names = F)
saveRDS(magnet_draws, "../saves/magnet_draws.rds")
```

```{r modelplot}
modelplot <- magnet_draws %>% 
  mutate(
    includes_zero = ifelse(predictor == "Age (scaled)", TRUE, FALSE)
  ) %>%
  ggplot(., aes(x = estimate, y = predictor, fill = includes_zero)) +
 tidybayes::stat_halfeye(.width = c(.95, .8), scale = 1, slab_color = "#006c66", slab_alpha = 0.5) +
    labs(x = "Posterior estimate when predicting TANGO mean", y = "") +
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("#006c66", "#EBECEC")) +
  coord_cartesian(expand = TRUE, xlim = c(-0.6, 0.4), ylim = c(1.5, 3.5)) +
  theme_classic()

modelplot

saveRDS(modelplot, file = "../saves/magnet_modelplot.rds")
# ggsave("../figures/modelplot.png", width = 10, height = 6, scale = 0.7, bg = "white")
```

```{r arrange_plot}
arrangedplot <- ggarrange(
  cor_ci, 
  modelplot, 
  ncol = 2, 
  labels = c("A", "B")
)

saveRDS(arrangedplot, file = "../saves/magnet_arrangedplot.rds")
ggsave("../figures/magnet_arrangedplot.png", width = 6, height = 2, scale = 1.75, bg = "white")
```
